\section{Negative Result Details}
\label{app:negative}

Section~6 and Figure~5 summarize four failed repairs.  This appendix
provides the corresponding experimental details, full data tables, and
additional ablations.  Unless otherwise noted, all repairs use the same
Track~A 100-pair PushT artifact (Appendix~\ref{app:track_a}) and the
default CEM hyperparameters: 300 samples, 30 elites, and 30 iterations.

\subsection{Learned Cost Heads}
\label{app:negative_cost_heads}

MLP cost heads $C_\psi(z,\zgoal)$ were trained on frozen LeWM latents using
V1 hinge labels.  Two data splits were used: Split~1 contains 15 easier
pairs with at least one successful action under default CEM, and Split~3
contains 16 hard pairs with zero successes under default CEM.  The split
numbering follows internal experiment IDs.

\begin{table}[h]
\centering
\caption{Endpoint ranking on Split~1 held-out pairs.}
\label{tab:h-cost-split1}
\small
\begin{tabular}{lcc}
\toprule
Cost function & Spearman & Pairwise accuracy \\
\midrule
Euclidean & 0.599 & 0.681 \\
Small $C_\psi$ & 0.442 & 0.777 \\
Large $C_\psi$ & 0.512 & 0.812 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]
\centering
\caption{Predicted endpoint metrics on Split~3 hard pairs across
cost-head variants.}
\label{tab:h-cost-split3}
\small
\begin{tabular}{lcc}
\toprule
Cost function & Spearman & Pairwise accuracy \\
\midrule
Euclidean & 0.233 & 0.578 \\
MLP offline small & 0.270 & 0.614 \\
MLP offline large & 0.352 & 0.611 \\
MLP mixed small & 0.285 & 0.594 \\
MLP mixed temp & 0.378 & 0.597 \\
Mahalanobis low-rank & 0.301 & 0.614 \\
CEM-aware & 0.227 & 0.579 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]
\centering
\caption{CEM planning success with learned cost heads.  No variant recovers
Split~3 hard-pair successes.}
\label{tab:h-cost-planning}
\small
\begin{tabular}{llcc}
\toprule
Split & Cost & Successes & Total \\
\midrule
Split~3 & Euclidean & 0 & 16 \\
Split~3 & Small $C_\psi$ & 0 & 16 \\
Split~3 & Large $C_\psi$ & 0 & 16 \\
Split~3 & Mixed $C_\psi$ & 0 & 16 \\
Split~3 & Mahalanobis & 0 & 16 \\
Split~3 & CEM-aware & 1 & 16 \\
Split~1 & Euclidean & 4 & 15 \\
Split~1 & $C_\psi$ & 3 & 15 \\
Split~1 & CEM-aware & 2 & 15 \\
\bottomrule
\end{tabular}
\end{table}

The mechanism is elite-cost compression.  At CEM iteration~30, the
Euclidean cost has elite standard deviation 0.354, while $C_\psi$ has elite
standard deviation 0.0015, a 238$\times$ compression.  This removes local
discrimination: all 30 elites receive nearly identical scores, so the
proposal update receives no gradient to separate near-successes from
near-misses, analogous to the V2 indicator degeneracy
(Appendix~\ref{app:oracle}).

\subsection{DINOv2 Encoder Replacement}
\label{app:negative_dinov2}

\begin{table}[h]
\centering
\caption{Endpoint ranking comparison between trained LeWM,
random-projected LeWM, and DINOv2 features on the Track~A 100-pair
artifact.}
\label{tab:h-dino}
\small
\begin{tabular}{lcccc}
\toprule
Representation & Dim & Spearman & Pairwise acc. & Per-pair $\rho$ \\
\midrule
LeWM (trained) & 192 & 0.503 & 0.647 & 0.355 \\
Random projection of LeWM & 192 & 0.501 & 0.643 & 0.349 \\
DINOv2 mean-pool & 768 & 0.261 & 0.610 & 0.285 \\
DINOv2 CLS & 768 & 0.239 & 0.594 & 0.248 \\
\bottomrule
\end{tabular}
\end{table}

DINOv2 mean-pool reaches the highest pairwise accuracy among non-LeWM
representations (0.610), but its Spearman correlation (0.261) is roughly
half of trained LeWM (0.503).  The random projection of LeWM latents
preserves 99.6\% of the Spearman signal (0.501 vs.\ 0.503), consistent
with Stage~1A (Appendix~\ref{app:stage1a}).  The 86.6M-parameter
ViT-B/14 of DINOv2 has 15.7$\times$ more parameters than the 5.5M-parameter
ViT-tiny, so the gap is not attributable to model capacity but to
task-specific training.

\subsection{Hybrid CEM Oracle Budget}
\label{app:negative_hybrid}

Hybrid CEM re-ranks the top $k$ CEM candidates per iteration using the V1
simulator cost from Equation~\ref{eq:v1}.  This tests whether injecting
ground-truth cost signal during search recovers hard-pair successes.

\begin{table}[h]
\centering
\caption{Hybrid CEM successes on Split~3 hard pairs by oracle budget.
Oracle budget is $k/N$ where $N=300$.}
\label{tab:h-oracle-split3}
\small
\begin{tabular}{cccc}
\toprule
$k$ & Oracle budget (\%) & Euclidean pre. & $C_\psi$ pre. \\
\midrule
0 & 0 & 0/16 & 0/16 \\
30 & 10 & 0/16 & 0/16 \\
60 & 20 & 7/16 & 6/16 \\
90 & 30 & 10/16 & 9/16 \\
150 & 50 & 12/16 & 14/16 \\
300 & 100 & 13/16 & 13/16 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]
\centering
\caption{Partial oracle schedules at $k=60$ on Split~3 hard pairs.  Only
sustained oracle access across all iterations recovers substantial
successes.}
\label{tab:h-oracle-partial}
\small
\begin{tabular}{lc}
\toprule
Schedule & Successes / 16 \\
\midrule
First iteration only & 0 \\
Early 5 iterations & 0 \\
Early 10 iterations & 2 \\
Every 3rd iteration & 4 \\
Late 10 iterations & 3 \\
All 30 iterations & 7 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]
\centering
\caption{Hybrid CEM successes on Split~1 easier pairs.  Euclidean prefilter
saturates at $k=150$.}
\label{tab:h-oracle-split1}
\small
\begin{tabular}{ccc}
\toprule
$k$ & Euclidean pre. & $C_\psi$ pre. ($k=150$) \\
\midrule
0 & 4/15 & -- \\
60 & 10/15 & -- \\
90 & 12/15 & -- \\
150 & 13/15 & 11/15 \\
300 & 13/15 & -- \\
\bottomrule
\end{tabular}
\end{table}

The 20\% oracle budget threshold ($k=60$) reported in Section~6.3
corresponds to 115{,}200 oracle rollouts and 2{,}880{,}000 simulator
environment steps over 30 CEM iterations.  This overhead makes hybrid CEM
impractical for deployment but confirms that the CEM cost landscape, not
action sampling, is the primary bottleneck.

\subsection{Subspace-CEM Stage A}
\label{app:negative_subspace}

Subspace-CEM searched with an $m=64$ projected cost and selected the rank-1
action by full 192-dimensional LeWM cost.  The locked Stage~A criterion
required at least +3 percentage points over the better of default CEM and
pure projected $m=64$.

\begin{table}[h]
\centering
\caption{Subspace-CEM Stage~A results over 30 pairs and two seeds.  Overall
$\Delta=-8.3$ percentage points versus default CEM.}
\label{tab:h-subspace}
\small
\begin{tabular}{lccc}
\toprule
Subset & Subspace (\%) & Default (\%) & Pure proj. $m{=}64$ (\%) \\
\midrule
Invisible quadrant & 0.0 & 0.0 & 6.25 \\
Ordinary & 75.0 & 83.3 & 85.0 \\
Latent favorable & 90.0 & 100.0 & 80.0 \\
V1 favorable & 20.0 & 40.0 & 20.0 \\
Overall & 48.3 & 56.7 & 50.0 \\
\bottomrule
\end{tabular}
\end{table}

Subspace-CEM regresses in every subset where default CEM has nonzero
success.  The largest regressions are in ordinary pairs ($-8.3$ percentage
points) and V1-favorable pairs ($-20.0$ percentage points).  The failure
shows that projected-cost search moves the CEM pool into a region that
full-dimensional re-ranking cannot recover, confirming that the
search-selection split identified in Section~4 cannot be repaired by a
simple two-stage approach.

\subsection{Cube Anomaly Multi-Seed Check}
\label{app:negative_cube_anomaly}

The Cube full projected CEM dimension sweep (Appendix~\ref{app:stage1b},
Section~G.3) showed a non-monotonic inverted-U success pattern peaking at
$m=32$.  A multi-seed sanity check added projection seeds~1 and~2 to the
existing seed~0 baseline.

\begin{table}[h]
\centering
\caption{Cube full projected CEM multi-seed check over 25 pairs.  The
$m=32$ versus $m=192$ gap is +5.33 percentage points across three seeds.}
\label{tab:h-cube-anomaly}
\small
\begin{tabular}{ccccc}
\toprule
$m$ & Seed 0 (\%) & Seed 1 (\%) & Seed 2 (\%) & 3-seed mean (\%) \\
\midrule
32 & 76.0 & 72.0 & 64.0 & 70.7 \\
192 & 60.0 & 64.0 & 72.0 & 65.3 \\
\bottomrule
\end{tabular}
\end{table}

The $m=32$ advantage over $m=192$ persists across seeds but with high
variance, including a seed-level range of 12 percentage points at $m=32$.
The closeout classified this as MIXED under the pre-registered
5--10 percentage-point criterion.  The final 50-pair, three-seed Cube
extension in Table~\ref{tab:cem-gap-protocols} supersedes the intermediate
$m=64$ sanity row from this check and reports 56.0\% rank-1 success.
