\documentclass{article}

\usepackage[preprint]{tmlr}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{array}
\usepackage{flafter}

\input{math_commands.tex}

\title{Endpoint Ranking Does Not Imply Planning: Auditing Cost Geometry in LeWM}

\author{
  Feng Ye \\
  \texttt{email@example.com}
}

% TMLR metadata
\def\month{05}
\def\year{2026}
\def\openreview{XXXX}

\newcommand{\zgoal}{z_g}
\newcommand{\zhatH}{\hat{z}_H}
\newcommand{\Cmodel}{C_\text{model}}
\newcommand{\Rendpoint}{R_\text{endpoint}}
\newcommand{\Rpool}{R_\text{pool}}
\newcommand{\DeltaCEM}{\Delta_\text{CEM}}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}

% REVISION: addresses W1 -- abstract and introduction scope narrowed.
% REVISION: addresses W2 -- mechanism-attribution subsection added in Section 4.
% REVISION: addresses W3 -- Cube full projected-CEM table updated with 50-pair, three-seed Phase E results.
% REVISION: addresses W4 -- MPPI cross-optimizer discussion added in Section 8.
% REVISION: addresses W5 -- privilege-free monitoring discussion added in Section 8.
% REVISION: addresses W1/W3/W4 -- limitations updated in Section 9.

\begin{document}

\maketitle

\begin{abstract}
LeWM-style latent planners that optimize Euclidean terminal latent costs with CEM can pass endpoint ranking while failing pool-level ranking. We audit this assumption in LeWM, a Joint-Embedding Predictive Architecture (JEPA)-style latent world model for goal-conditioned control from pixels, and find that endpoint ranking and CEM final-pool ranking can decouple. At projection dimension $m=64$, the CEM Compatibility Gap $\DeltaCEM$ reaches 0.43 on PushT and 0.57 on OGBench-Cube, while rank-1 planning success is 29.6\% and 56.0\%, respectively. The audit covers two environments, four protocol settings, and eight random-geometry controls with fixed artifacts and pre-registered decision rules. Four simple repairs fail under their pre-specified thresholds. A learned post-hoc calibration experiment shows that pool-level ranking labels carry strong corrective signal on seen pools, but the local geometry distortion is pair-specific and does not transfer to held-out configurations, pointing to training-time or pool-conditioned approaches as a plausible path forward. These results localize the problem to the interaction between cost geometry and CEM-local candidate pools. A cross-architecture check on DINO-WM reveals analogous pool-level cost collapse, though DINO-WM's near-zero endpoint ranking prevents it from serving as a clean replication of the endpoint-pool decoupling pattern. % REVISION: addresses W1
For LeWM-style latent planners that select actions by terminal latent costs under iterative sampling-based optimization, we recommend reporting pool-level and selection-level metrics alongside endpoint metrics.
\end{abstract}

\input{sections/introduction}
\input{sections/background}
\input{sections/framework}
\input{sections/decoupling}
\input{sections/nulls}
\input{sections/negative}
\input{sections/warp_ensemble}
\input{sections/discussion}
\input{sections/limitations}

\section*{Ethics Statement}
This work audits an existing publicly released model and does not involve human subjects, private data, or deployment of autonomous systems.

\section*{Reproducibility Statement}
All experiments use publicly available model checkpoints and datasets. The three-level diagnostic protocol requires only saving CEM final candidate pools and scoring them with an evaluation-time physical cost. Code, frozen JSON artifacts, and pool data will be released upon publication. Appendix~\ref{app:reproducibility} details the computational environment, artifact inventory, and checkpoint provenance.

\bibliographystyle{tmlr}
\bibliography{references}

\newpage
\appendix

\input{sections/appendix_a_reproducibility}
\input{sections/appendix_b_phase0}
\input{sections/appendix_c_track_a}
\input{sections/appendix_d_oracle}
\input{sections/appendix_e_stage1a}
\input{sections/appendix_f_c6_audit}
\input{sections/appendix_g_stage1b}
\input{sections/appendix_h_negative}
\input{sections/appendix_i_preregistration}

\end{document}
