\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{array}
\usepackage{hyperref}
\usepackage{xcolor}

\newcommand{\Rendpoint}{R_{\mathrm{endpoint}}}
\newcommand{\Rpool}{R_{\mathrm{pool}}}
\newcommand{\DeltaCEM}{\Delta_{\mathrm{CEM}}}
\newcommand{\Cmodel}{C_{\mathrm{model}}}
\newcommand{\CVone}{C_{\mathrm{V1}}}
\newcommand{\CVthree}{C_{\mathrm{V3}}}
\newcommand{\Creal}{C_{\mathrm{real\_state}}}

\title{Response to Reviewer Comments}
\author{}
\date{}

\begin{document}
\maketitle

\section*{Response to Second-Round Reviewer Comments}

We thank the reviewer for the detailed and constructive second-round feedback (6/10). Below we address each of the six new concerns.

\subsection*{New W1: Generalization scope still too broad}

\paragraph{Concern.}
The paper's claims still read as general conclusions about latent world models.

\paragraph{Response.}
We have further narrowed scope language. The abstract now reads ``For LeWM-style latent planners...'' and the Cube $m=64$ success has been corrected from 72.0\% to 56.0\% to match the extended 50-pair Table 2 data. We position the paper as a rigorous LeWM-specific audit that proposes a reusable diagnostic protocol, not a universal empirical law.

\subsection*{New W2: V1 oracle cannot alone complete mechanism attribution}

\paragraph{Concern.}
V1 uses physical variables directly. It cannot distinguish learned representation failure from predictor rollout error or cost shape issues. A V3 (actual-terminal encoder latent cost) comparison on the same pool is needed.

\paragraph{Response.}
We conducted the requested V3 same-pool ranker decomposition on all 100 PushT final pools. For each of the 300 candidates per pair, we replayed the raw actions in the simulator, encoded the actual terminal observation with the LeWM encoder, and computed $\CVthree=\|z_\mathrm{terminal}-z_g\|_2^2$.

\begin{center}
\begin{tabular}{lcc}
\toprule
Cost ranker & $\Rpool$ (mean $\pm$ std) & 95\% bootstrap CI \\
\midrule
$\Cmodel$ (predicted terminal latent) & $0.084 \pm 0.243$ & $[0.037, 0.131]$ \\
$\CVthree$ (actual terminal latent) & $0.258 \pm 0.566$ & $[0.146, 0.365]$ \\
$\CVone$ (physical hinge oracle) & $0.666 \pm 0.357$ & $[0.595, 0.735]$ \\
\bottomrule
\end{tabular}
\end{center}

The decomposition reveals a three-level attribution. Removing predictor error ($\Cmodel \rightarrow \CVthree$) improves $\Rpool$ by 0.174, confirming that predictor rollout degrades pool ranking. Removing both predictor and encoder geometry ($\CVthree \rightarrow \CVone$) improves $\Rpool$ by a further 0.408. Even with perfect prediction, the encoder's Euclidean latent distance remains a weak pool ranker.

This directly addresses the reviewer's concern: the problem is not solely predictor error, and the encoder's terminal-latent geometry / Euclidean cost shape is the dominant bottleneck. We have revised Section 4.4 to include this decomposition table.

\paragraph{Location.}
Section 4.4, new decomposition table and revised interpretation.

\subsection*{New W3: Statistical uncertainty insufficient}

\paragraph{Concern.}
Key claims like ``22\% CEM-specific'' and ``78\% representation-driven'' lack confidence intervals.

\paragraph{Response.}
We computed bootstrap 95\% CIs (10,000 resamples) for all core metrics:
\begin{itemize}
  \item $\DeltaCEM = 0.422$ with CI $[0.374, 0.468]$.
  \item $\Rpool(\Cmodel) = 0.084$ with CI $[0.037, 0.131]$.
  \item $\Rpool(\CVone) = 0.666$ with CI $[0.595, 0.735]$.
  \item MPPI attribution = 22\% with CI $[-16\%, 48\%]$; the point estimate is positive but the CI spans zero.
\end{itemize}

We have revised the MPPI discussion to present the attribution as a point estimate with acknowledged uncertainty rather than a precise decomposition. The qualitative finding -- that soft weighting improves but does not eliminate the endpoint-pool gap -- is robust regardless of the exact percentage.

For repair failures, we added Clopper--Pearson exact binomial CIs (e.g., cost head: 0/16, 95\% upper bound = 20.6\%).

\paragraph{Location.}
Core metrics in Sections 4.4, 7, and 8 now include bootstrap CIs or exact intervals where applicable.

\subsection*{New W4: Cube is Case B, not planning failure}

\paragraph{Concern.}
Cube has $\Rpool \approx 0$ but high success, fitting Case B (pool already good, ranking less important) rather than planning failure.

\paragraph{Response.}
We agree and have added an explicit sentence in Section 4.2: ``In Cube, the near-zero $\Rpool$ coexists with relatively high planning success, consistent with Case B in our taxonomy: the pool contains sufficient success mass that ranking is less critical for action selection. Cube therefore provides cross-environment evidence for the endpoint-pool ranking gap, not for cost-driven planning failure.''

\paragraph{Location.}
Section 4.2, after Table 2 discussion.

\subsection*{New W5: Text consistency (72.0\% vs 56.0\%, missing Section 7)}

\paragraph{Concern.}
Abstract Cube number conflicts with Table 2; Section 7 is missing.

\paragraph{Response.}
Fixed. The abstract now reports 56.0\% matching the extended 50-pair Table 2 data. The section numbering gap has been removed (now flows Section 6 $\rightarrow$ Section 7 calibration experiment $\rightarrow$ Section 8 Discussion $\rightarrow$ Section 9 Limitations). The 72.0\% figure from the original 25-pair single-seed run remains only in the appendix seed-level sanity table, clearly labeled as preliminary.

\paragraph{Location.}
Abstract, section headers throughout.

\subsection*{New W6: Warp Ensemble Selection is not a strict held-out evaluation}

\paragraph{Concern.}
In the original Warp Ensemble Selection experiment, each test pool contributed labels to 9 of the 10 ensemble members through the 10-fold setup. The reported gain could therefore reflect training exposure rather than genuine generalization.

\paragraph{Response.}
We agree. We ran a strict held-out evaluation that separates calibration capacity from generalization. Pairs 0--69 are used for training only. Within those 70 pairs, we train ten warps by ten-fold cross-validation. Pairs 70--99 are then evaluated as a completely held-out test set that is never seen by any warp during training.

This strict split changes the conclusion of Section~7. On seen pools, the calibration still learns strong corrective structure: mean $\Rpool$ rises from 0.086 under Euclidean ranking to 0.441 when each training pool is evaluated by the nine warps that saw it during training. Thus pool-level ranking labels do contain strong local corrective signal.

However, the learned calibration does not transfer to unseen pairs. On the 30 strictly held-out pools, Euclidean selection and warp-ensemble rank averaging both achieve 13.3\% success; the exact paired McNemar test gives $p=1.0$. Mean regret is slightly worse for the warp ensemble (37.67 versus 36.96). We also repeated the matched planning setting with default CEM search plus post-hoc warp reranking on the same held-out 30 pairs with three seeds. Default CEM succeeds on 20.0\% of runs, while warp reranking reaches 18.9\%; McNemar again gives $p=1.0$.

We therefore rewrote Section~7 as a diagnostic validation experiment rather than a method contribution. The revised section now emphasizes the scientific conclusion: pool-level calibration works on seen pools but fails to generalize, implying that the local geometry distortion is pair-specific and that training-time representation intervention is likely necessary.

\paragraph{Location.}
Rewritten Section 7, new held-out tables and interpretation.

\clearpage

\section*{Response to First-Round Reviewer Comments}

We thank the reviewer for the thorough and constructive evaluation. Below we address each weakness with new experiments and text revisions. All supplementary experiments, scripts, and data are released in the repository under \path{scripts/revision/}, \path{results/revision/}, and \path{docs/revision/}.

\section*{W1: Single-model, dual-environment generalization}

\paragraph{Reviewer concern.}
All experiments use only LeWM on PushT and OGBench-Cube. The abstract and introduction overclaim generality with phrases like ``any latent world model that plans with terminal latent costs can be audited this way.''

\paragraph{Our response.}
We agree that the empirical evidence is bounded to one model and two environments, and have revised all overclaiming language accordingly.

\textit{Abstract:} The original sentence has been replaced with: ``For LeWM-style latent planners that select actions by terminal latent costs under iterative sampling-based optimization, we recommend reporting pool-level and selection-level metrics alongside endpoint metrics.''

\textit{Introduction (Section 1):} We now state explicitly: ``Our empirical claims apply to LeWM with Euclidean terminal latent costs under CEM-style optimization, evaluated on PushT and OGBench-Cube. The evaluation principle is broader, but it should be applied with the same instrumentation.''

\textit{Limitations (Section 9):} Three explicit scope boundaries are now listed: (1) single model, two environments; (2) CEM as primary optimizer with partial MPPI cross-check; (3) privileged physical cost required for full evaluation.

\textit{Cross-model feasibility:} We investigated PLDM (Sobal et al., 2025) as a cross-model validation target. The public repository exists but no ready-to-use PushT checkpoint was available at revision time. We leave cross-model validation as future work, noting that the diagnostic protocol ($\Rendpoint$, $\Rpool$, $\DeltaCEM$) is model-agnostic and directly applicable to any latent world model that exposes its candidate pool.

\paragraph{Location in revised paper.}
Abstract (page 1), Section 1 (page 2, lines 88--97), Section 9 (Limitations).

\section*{W2: Diagnostic power of $\DeltaCEM$ -- attribution of $\Rpool \approx 0$}

\paragraph{Reviewer concern.}
$\Rpool \approx 0$ could reflect multiple mechanisms: (a) CEM convergence compresses candidates into a physically narrow neighborhood where any cost fails to discriminate, (b) the encoder genuinely loses local order preservation, or (c) predictor bias. The paper does not experimentally distinguish these.

\paragraph{Our response.}
We conducted a new Rpool(V1) attribution experiment on all 100 PushT default CEM final pools. For each pair, we computed the Spearman correlation of the V1 oracle hinge cost (a privileged physical cost) against $\Creal$ within the same 300-candidate pool, and compared it with the learned Euclidean cost.

\textit{Key results:}

\begin{center}
\begin{tabular}{lrrr}
\toprule
Subset & $n$ & $\Rpool(\Cmodel)$ & $\Rpool(\CVone)$ \\
\midrule
Overall & 100 & 0.084 & 0.666 \\
Invisible quadrant & 16 & 0.022 & 0.775 \\
Sign reversal & 21 & $-0.040$ & 0.925 \\
Latent-favorable & 12 & 0.197 & 0.417 \\
Ordinary & 47 & 0.146 & 0.581 \\
\bottomrule
\end{tabular}
\end{center}

The V1 oracle retains strong pool-level ranking ($\Rpool(\CVone) = 0.666$) where the learned cost fails ($\Rpool(\Cmodel) = 0.084$). This rules out pure CEM convergence compression as the dominant explanation: if the pool were compressed into a physically indistinguishable neighborhood, V1 would also fail. Instead, the data clearly indicates \textbf{local representation failure}: the learned Euclidean cost loses order-preservation in the CEM-converged region while the privileged oracle cost does not.

The effect is strongest in the invisible quadrant (16 pairs where endpoint metrics look healthy but planning fails): $\Rpool(\CVone) = 0.775$ while $\Rpool(\Cmodel) = 0.022$, with pool success mass of only 1.4\%. The CEM pool contains successful candidates, but the learned cost cannot identify them.

\paragraph{Location in revised paper.}
New Section 4.4, ``Mechanism attribution: convergence compression versus local representation failure'' (page 8).

\section*{W3: Weak Cube evidence}

\paragraph{Reviewer concern.}
Cube full projected CEM uses only 25 pairs $\times$ 1 seed (vs. 100 pairs $\times$ 3 seeds for re-rank-only). The inverted-U pattern received MIXED status. This asymmetry weakens cross-environment claims.

\paragraph{Our response.}
We extended the Cube full projected CEM experiment from 25 pairs $\times$ 1 seed to \textbf{50 pairs $\times$ 3 projection seeds} across all five projection dimensions ($m \in \{1, 8, 32, 64, 192\}$), totaling 750 CEM rollouts with full 300-candidate simulator scoring ($\sim$10 hours on Apple Silicon MPS).

\textit{Updated results:}

\begin{center}
\begin{tabular}{rrrr}
\toprule
$m$ & Full-CEM success (50p, 3s) & Re-rank success (100p, 3s) & Gap \\
\midrule
1 & $32.7\% \pm 8.1\%$ & $42.7\% \pm 3.2\%$ & $-10.0$ pp \\
8 & $52.7\% \pm 4.2\%$ & $43.3\% \pm 1.5\%$ & $+9.3$ pp \\
32 & $61.3\% \pm 1.2\%$ & $46.3\% \pm 2.5\%$ & $+15.0$ pp \\
64 & $56.0\% \pm 5.3\%$ & $49.7\% \pm 2.5\%$ & $+6.3$ pp \\
192 & $62.0\% \pm 6.0\%$ & $47.7\% \pm 1.2\%$ & $+14.3$ pp \\
\bottomrule
\end{tabular}
\end{center}

$\Rpool(\Cmodel)$ across all dimensions ranges from $-0.001$ to 0.023, consistently near zero, confirming endpoint-planning decoupling in Cube at scale.

The original inverted-U success pattern (peak at $m=32$, decline at $m=192$) \textbf{does not persist} at 50-pair, 3-seed scale. Success plateaus from $m=32$ onward, indicating the original pattern was small-sample noise. This simplifies the cross-environment comparison: both PushT and Cube show rising $\Rendpoint$ with dimension while $\Rpool$ remains near zero.

Table 2 in the revised paper now uses the extended 50-pair $\times$ 3-seed Cube data.

\paragraph{Location in revised paper.}
Table 2 (page 7), Section 4.2 discussion (page 7, lines 368--377), Section 8 Limitations (page 12).

\section*{W4: No alternative optimizer comparison}

\paragraph{Reviewer concern.}
The paper does not test whether endpoint-planning decoupling is CEM-specific (due to hard elite truncation) or generalizes to other iterative optimizers like MPPI.

\paragraph{Our response.}
We implemented an MPPI planner that replaces CEM's hard elite selection with softmax-weighted averaging over all 300 candidates, keeping all other parameters identical (same model, action parameterization, horizon, 30 iterations, 300 candidates). After a temperature sweep ($\tau \in \{0.01, 0.1, 1.0, 10.0\}$), we selected $\tau^* = 1.0$ and ran the experiment on 30 PushT pairs $\times$ 3 seeds (90 scored pools).

\textit{Comparison:}

\begin{center}
\begin{tabular}{lrr}
\toprule
Metric & CEM default & MPPI ($\tau=1.0$) \\
\midrule
Planning success & 53.3\% & 38.9\% \\
$\Rpool(\Cmodel)$ & 0.115 & 0.194 \\
$\Rpool(\CVone)$ & 0.537 & 0.707 \\
Pool $\Creal$ std & 7.947 & 29.478 \\
\bottomrule
\end{tabular}
\end{center}

MPPI preserves \textbf{3.7$\times$ more physical diversity} in its final pool than CEM (std 29.478 vs. 7.947), confirming that CEM's hard truncation compresses pool diversity. Correspondingly, $\Rpool(\Cmodel)$ improves from 0.115 to 0.194 under MPPI.

However, MPPI still retains only $\sim$39\% of the endpoint ranking signal ($\Rpool/\Rendpoint = 0.194/0.495$). Using $\Rendpoint = 0.495$ at $m=192$ as the reference, the CEM-to-MPPI improvement has a 22\% point estimate of the endpoint-pool gap, but the paired-bootstrap 95\% confidence interval is wide ($[-16\%,48\%]$). Thus endpoint-planning decoupling is \textbf{partially optimizer-specific} (CEM's hard truncation worsens pool geometry) and \textbf{partially optimizer-general}: the majority of the decoupling appears representation-driven, with the MPPI improvement modest relative to the full endpoint-pool gap.

The lower MPPI planning success (38.9\% vs. 53.3\%) is consistent with slower convergence under 30 iterations and does not invalidate the pool-level comparison, which is the relevant diagnostic.

\paragraph{Location in revised paper.}
Section 8.3, ``Optimizer dependence: CEM versus MPPI.''

\section*{W5: Privileged physical cost requirement limits practical applicability}

\paragraph{Reviewer concern.}
All pool-level metrics require $\Creal$ (simulator access). How can one detect $\DeltaCEM$ in deployment without a simulator?

\paragraph{Our response.}
We tested whether privilege-free proxy metrics (computable from the learned cost alone, no simulator needed) correlate with selection regret across the 100 PushT pairs.

\textit{Results:}

\begin{center}
\begin{tabular}{lrr}
\toprule
Proxy metric & Corr. with selection regret & $p$-value \\
\midrule
\texttt{top30\_Cmodel\_std} & 0.314 & 0.001 \\
\texttt{pool\_Cmodel\_std} & 0.254 & 0.011 \\
\texttt{C\_model\_dynamic\_range} & 0.223 & 0.026 \\
\bottomrule
\end{tabular}
\end{center}

Within the invisible-quadrant subset, the signal is stronger: \texttt{pool\_Cmodel\_std} vs. selection regret has $\rho = 0.588$ ($p = 0.017$). As a cross-check, \texttt{pool\_Cmodel\_std} tracks \texttt{pool\_Creal\_std} at $\rho = 0.482$ ($p < 0.001$), confirming that learned-cost spread partially reflects physical spread.

No single global proxy crosses our pre-specified $\rho > 0.4$ threshold, so these proxies are not full replacements for the privileged audit. However, they demonstrate that \textbf{partial monitoring is feasible without simulator access}: low learned-cost diversity in the CEM final pool can flag candidate-pool compression and elevated selection risk.

\paragraph{Location in revised paper.}
Section 8.1, paragraph on privilege-free monitoring.

\section*{Summary of Changes}

\begin{center}
\footnotesize
\begin{tabular}{@{}>{\raggedright\arraybackslash}p{0.09\linewidth}>{\raggedright\arraybackslash}p{0.27\linewidth}>{\raggedright\arraybackslash}p{0.33\linewidth}>{\raggedright\arraybackslash}p{0.16\linewidth}@{}}
\toprule
Weakness & Experiment & Key finding & Paper location \\
\midrule
W1 & Text revision + PLDM feasibility & Scope narrowed; PLDM deferred & Abstract, Sec. 1, Sec. 8 \\
W2 & Rpool(V1) attribution (100 pairs) & Local representation failure dominant & New Sec. 4.4 \\
W3 & Cube extended (50p $\times$ 3s, 750 runs) & Decoupling robust at scale; inverted-U gone & Table 2, Sec. 4.2 \\
W4 & MPPI comparison (30p $\times$ 3s, 90 pools) & Soft weighting improves but does not close the gap; attribution CI is wide & Sec. 8.3 \\
W5 & Privilege-free proxy (100 pairs) & Partial monitoring feasible ($\rho$ up to 0.588) & Sec. 8.1 \\
W6 & Strict held-out WES evaluation & Seen-pool calibration is strong, but held-out generalization fails & Rewritten Sec. 7 \\
\bottomrule
\end{tabular}
\end{center}

All supplementary scripts, results, and memos are available in the repository under \path{scripts/revision/}, \path{results/revision/}, and \path{docs/revision/}. The progress document \path{docs/revision/revision_progress.md} provides a complete timeline and artifact inventory.

\end{document}
